---
title: "Reproduction of Winter & Bergen 2012 using LMEs"
author: "Cameron Jones"
date: "4/12/2023"
output:
  html_document: 
    toc: yes
    toc_float: yes
    theme: flatly
    highlight: kate
    code_folding: hide
    number_sections: yes
    fig_caption: yes
---

# Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Import files and libraries

```{r}
library(tidyverse)
library(lmerTest)

e1.raw <- read.csv("../data/winter2012/e1/human_data.csv")
e2.raw <- read.csv("../data/winter2012/e2/human_data.csv")

```

# Experiment 1

## Preprocessing

```{r}

e1 <- e1.raw %>%
  filter(
    CRITICAL == "CRITICAL"
  ) %>%
  mutate(
    item_type = ifelse(is.na(LM), "EXPL", "LM"),
    match = recode(SAME, "SAME"="Match", "DIFF"="Mismatch"),
    item_id = paste(item_type, ITEM, sep="_")
  ) %>%
  separate(
    CONDITION, c("sentence_type", "image_type")
  ) %>%
  rename(
    ppt_id = Subject,
    block = Block,
    item = ITEM,
    accuracy = Pic.ACC,
    rt = Pic.RT
  ) %>%
  select(
    ppt_id, block, item_type, item_id, match, sentence_type, image_type, accuracy, rt
  )

```

## Descriptive

22 ppts did 32 items each

```{r}

e1 %>%
  group_by(ppt_id) %>%
  summarize(n_items=n()) %>%
  group_by(n_items) %>%
  summarize(n_ppts=n())


```

## Preprocess

```{r}

e1.clean <- e1 %>%
  filter(
    accuracy == 1,
    rt > 300,
    rt < 3000
  )

```

## Mean RT per match condition

```{r}

e1.clean %>%
  group_by(match) %>%
  summarize(mean(rt))

```

## ANOVA

### By subject

```{r}

summary(aov(rt ~ sentence_type * image_type + Error(factor(ppt_id)), data=e1.clean))

```


### By item

```{r}

summary(aov(rt ~ sentence_type * image_type + Error(factor(item_id)), data=e1.clean))

```

## Linear Model

### Match effect

```{r}


m.e1.base <- lmer(rt ~ 1 + (1  | ppt_id) + (1  | item_id),
                  data=e1.clean, REML=FALSE)

m.e1 <- lmer(rt ~ match + (1 | ppt_id) + (1 | item_id),
                  data=e1.clean, REML=FALSE)

anova(m.e1, m.e1.base)

summary(m.e1)


```

### Interaction


```{r}


m.e1.ixn.base <- lmer(rt ~ 1 + sentence_type + image_type + 
                        (1 | ppt_id) + (1 | item_id),
                  data=e1.clean, REML=FALSE)

m.e1.ixn <- lmer(rt ~ sentence_type * image_type +
                   (1 | ppt_id) + (1 | item_id),
                  data=e1.clean, REML=FALSE)

anova(m.e1.ixn, m.e1.ixn.base)

summary(m.e1.ixn)


```
# Experiment 2

## Preprocessing

```{r}

e2.ppt.ex.acc <- e2.raw %>%
  group_by(Subject) %>%
  summarize(acc = mean(SoundOut1.ACC)) %>%
  arrange(acc) %>%
  filter(acc < 0.8) %>%
  pull(Subject)

unique(e2.raw$FirstLanguage)

```

```{r}

e2 <- e2.raw %>%
  filter(
    Type == "CRITICAL",
    !Subject %in% e2.ppt.ex.acc,
    FirstLanguage %in% c("English", "english")
  ) %>%
  mutate(
    match = recode(SAME, "SAME"="Match", "DIFF"="Mismatch"),
    item_id = ITEM
  ) %>%
  separate(
    CONDITION, c("sentence_type", "sound_type")
  ) %>%
  rename(
    ppt_id = Subject,
    block = Block,
    item = ITEM,
    accuracy = SoundOut1.ACC,
    rt = SoundOut1.RT
  ) %>%
  select(
    ppt_id, block, item_id, match, sentence_type, sound_type, accuracy, rt
  )

```

## Descriptive

32 ppts did 24 items each (after exclusions for language & sound verification accuracy)

```{r}

e2 %>%
  group_by(ppt_id) %>%
  summarize(n_items=n()) %>%
  group_by(n_items) %>%
  summarize(n_ppts=n())


```

## Preprocess

```{r}

e2.clean <- e2 %>%
  filter(
    accuracy == 1,
    rt > 300,
    rt < 3000
  )

```

## Mean RT per match condition

```{r}

e2.clean %>%
  group_by(match) %>%
  summarize(mean(rt))

```

## ANOVA

### By subject

```{r}

summary(aov(rt ~ sentence_type * sound_type + Error(factor(ppt_id)), data=e2.clean))

```


### By item

```{r}

summary(aov(rt ~ sentence_type * sound_type + Error(factor(item_id)), data=e2.clean))

```

## Linear Model

### Match effect

```{r}


m.e2.base <- lmer(rt ~ 1 + (1  | ppt_id) + (1  | item_id),
                  data=e2.clean, REML=FALSE)

m.e2 <- lmer(rt ~ match + (1 | ppt_id) + (1 | item_id),
                  data=e2.clean, REML=FALSE)

anova(m.e2, m.e2.base)

summary(m.e2)


```

### Interaction

```{r}


m.e2.ixn.base <- lmer(rt ~ 1 + sentence_type + sound_type + 
                        (1 | ppt_id) + (1 | item_id),
                  data=e2.clean, REML=FALSE)

m.e2.ixn <- lmer(rt ~ sentence_type * sound_type +
                   (1 | ppt_id) + (1 | item_id),
                  data=e2.clean, REML=FALSE)

anova(m.e2.ixn, m.e2.ixn.base)

summary(m.e2.ixn)


```


