---
title: "Reproduction of Zwaan & Pecher 2012 using LMEs"
author: "Anonymous"
date: "29/11/2023"
output:
  html_document: 
    toc: yes
    toc_float: yes
    theme: flatly
    highlight: kate
    code_folding: hide
    number_sections: yes
    fig_caption: yes
---

# Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Import files and libraries

```{r}
library(tidyverse)
library(lmerTest)

e1a <- read.csv("../data/pecher2012/human_data/Experiment_1a.csv")
e1b <- read.csv("../data/pecher2012/human_data/Experiment_1b.csv")
e2a <- read.csv("../data/pecher2012/human_data/Experiment_2a.csv")
e2b <- read.csv("../data/pecher2012/human_data/Experiment_2b.csv")

```


# Experiment 1A

## Descriptive

No ppts

```{r}

n_distinct(e1a$ppt_id)

```

Median rt matches values in paper

```{r}

e1a.correct <- e1a %>%
  filter(accuracy == 1)

e1a.correct %>%
  group_by(ppt_id, match) %>%
  summarize(
    rt = median(rt)
  ) %>%
  group_by(match) %>%
  summarize(
    rt = mean(rt)
  )
  

```

## ANOVA

```{r}

anova_data <- e1a.correct %>%
  group_by(ppt_id, match) %>%
  summarize(median_rt = median(rt)) %>%
  mutate(
    ppt_id = factor(ppt_id),
    match = factor(match)
  ) %>%
  ungroup()

summary(aov(median_rt ~ match + Error(factor(ppt_id)), data=anova_data))

```

## Linear Model

```{r}

e1a.clean <- e1a.correct %>%
  filter(
    rt > 0.3,
    rt < 3
  )

m.e1a.base <- lmer(rt ~ 1 + (1  | ppt_id) + (1  | object),
                  data=e1a.clean, REML=FALSE)

m.e1a <- lmer(rt ~ match + (1  | ppt_id) + (1  | object),
                  data=e1a.clean, REML=FALSE)

anova(m.e1a, m.e1a.base)

summary(m.e1a)

```


# Experiment 1B

## Descriptive

Note, 1 fewer than ZP, appears to be due to language filtering

```{r}

n_distinct(e1b$ppt_id)

```

Median rt matches values in paper

```{r}

e1b.correct <- e1b %>%
  filter(accuracy == 1)

e1b.correct %>%
  group_by(ppt_id, match) %>%
  summarize(
    rt = median(rt)
  ) %>%
  group_by(match) %>%
  summarize(
    rt = mean(rt)
  )
  

```

## ANOVA

```{r}

anova_data <- e1b.correct %>%
  group_by(ppt_id, match) %>%
  summarize(median_rt = median(rt)) %>%
  mutate(
    ppt_id = factor(ppt_id),
    match = factor(match)
  ) %>%
  ungroup()

summary(aov(median_rt ~ match + Error(factor(ppt_id)), data=anova_data))


```

## Linear Model

```{r}

e1b.clean <- e1b.correct %>%
  filter(
    rt > 0.3,
    rt < 3
  )

m.e1b.base <- lmer(rt ~ 1 + (1   | ppt_id) + (1  | object),
                  data=e1b.clean, REML=FALSE)

m.e1b <- lmer(rt ~ match + (1   | ppt_id) + (1  | object),
                  data=e1b.clean, REML=FALSE)

anova(m.e1b, m.e1b.base)

summary(m.e1b)

```


# Experiment 2A

## Descriptive

No ppts

```{r}

n_distinct(e2a$ppt_id)

```

Median rt matches values in paper

```{r}

e2a.correct <- e2a %>%
  filter(accuracy == 1)

e2a.correct %>%
  group_by(ppt_id, match) %>%
  summarize(
    rt = median(rt)
  ) %>%
  group_by(match) %>%
  summarize(
    rt = mean(rt)
  )
  

```

## ANOVA

```{r}

anova_data <- e2a.correct %>%
  group_by(ppt_id, match) %>%
  summarize(median_rt = median(rt)) %>%
  mutate(
    ppt_id = factor(ppt_id),
    match = factor(match)
  ) %>%
  ungroup()

summary(aov(median_rt ~ match + Error(factor(ppt_id)), data=anova_data))

```

## Linear Model

```{r}

e2a.clean <- e2a.correct %>%
  filter(
    rt > 0.3,
    rt < 3
  )

m.e2a.base <- lmer(rt ~ 1 + (1  | ppt_id) + (1  | object),
                  data=e2a.clean, REML=FALSE)

m.e2a <- lmer(rt ~ match + (1  | ppt_id) + (1  | object),
                  data=e2a.clean, REML=FALSE)

anova(m.e2a, m.e2a.base)

summary(m.e2a)

```


# Experiment 2B

## Descriptive

No ppts

Note: 1 additional ppt, 

```{r}

n_distinct(e2b$ppt_id)

```

Median rt matches values in paper

```{r}

e2b.correct <- e2b %>%
  filter(accuracy == 1)

e2b.correct %>%
  group_by(ppt_id, match) %>%
  summarize(
    rt = median(rt)
  ) %>%
  group_by(match) %>%
  summarize(
    rt = mean(rt)
  )
  

```

## ANOVA

```{r}

anova_data <- e2b.correct %>%
  group_by(ppt_id, match) %>%
  summarize(median_rt = median(rt)) %>%
  mutate(
    ppt_id = factor(ppt_id),
    match = factor(match)
  ) %>%
  ungroup()

summary(aov(median_rt ~ match + Error(factor(ppt_id)), data=anova_data))

```

## Linear Model

```{r}

e2b.clean <- e2b.correct %>%
  filter(
    rt > 0.3,
    rt < 3
  )

m.e2b.base <- lmer(rt ~ 1 + (1  | ppt_id) + (1  | object),
                  data=e2b.clean, REML=FALSE)

m.e2b <- lmer(rt ~ match + (1  | ppt_id) + (1  | object),
                  data=e2b.clean, REML=FALSE)

anova(m.e2b, m.e2b.base)

summary(m.e2b)

```

